Containers and R: the Rockerverse and beyond

### 2019-10-25 manuscript.Rmd
Add section on shared work environments and GPUs

- Binder
  - [`holepunch`](https://github.com/karthik/holepunch) streamlines making an R research compendium Binder-ready [@karthik]
  - [`repo2docker`](https://repo2docker.readthedocs.io/en/latest/config_files.html) (`install.R`, `DESCRPTION`, `runtime.txt`) based on the Rocker images [@nuest]
- GPU [@noamross, @cboettig]
- Gigantum stack
- in education (ready to use teaching/learning environments)
  - [`r-db`](https://github.com/ColinFay/r-db) [@ColinFay]
- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?

>>>>

- Binder
  - [`holepunch`](https://github.com/karthik/holepunch) streamlines making an R research compendium Binder-ready [@karthik]
  - [`repo2docker`](https://repo2docker.readthedocs.io/en/latest/config_files.html) (`install.R`, `DESCRPTION`, `runtime.txt`) based on the Rocker images [@nuest]
- GPU [@noamross, @cboettig]

One use for containers is to run on shared local hardware where teams manage
their own high-performance servers  This can follow one of several
design patterns: users may deploy containers to hardware as a work environment
for a specific project, conatiners may provide per-user persistent environments, or a single container can act as
a common multi-user environment for a server.  The former models provide modularity,
while the latter approach is most similar to a simple shared server. In all cases,
though, the containerized approach provides several advantages: First, users may
use the same image and thus work environment on desktop and laptop computers, as well.
Second, software updates can be achieved by updating and redeploying the container,
rather than tracking local installs on each server.  Third, the containerized
environment can be quickly deployed to other hardware, cloud or local, if 
more resources are neccessary or in case of server destruction or failure. In any of
these cases, users need a method to interact with the containers, be it and
IDE, or command-like access and tools such as SSH, which is usually not part
of standard container recipes and must be added.  The Rocker project provides
containers pre-installed with the RStudio IDE.

In cases where users store nontrivial amounts of data for their projects, data
needs to persist beyond the life of the container. This may be via in shared disks, attached
network volumes, or in separate storage where it is uploaded between sessions. In
the case of shared disks or network-attached volumes, care must be taken to
persist user permissions, and of course backups are still neccessary.
When working with multiple servers, an automation
framework such as [Ansible](https://www.ansible.com) may be useful for managing
users, permisions, and disks along with containers.

In some cases, containers are not completely portable between hardware environments.
This is the case for software using graphical processing units (GPUs)
which are increasingly popular for compute-intensive machine learning tasks.
Containers running GPU software  require drivers and libraries specific
to GPU models and versions, and containers require a specialized runtime to connect
to the underlying GPU hardware. For NVIDIA GPUs, the [NVIDIA Container Toolkit](https://github.com/NVIDIA/nvidia-docker)
includes a specialized runtime plugn for docker and set of base images with
appropriate drivers and libraries, and the Rocker project has (beta) images
based on these that include GPU-enabled versions of machine-learning R packages.

- Gigantum stack
- in education (ready to use teaching/learning environments)
  - [`r-db`](https://github.com/ColinFay/r-db) [@ColinFay]
- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?
